News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-07-16 rupnik2015.tex
Bib and main paper fixes

We also used a term re-weighting scheme that adjusts for the fact that some words occur more frequently in general. A term weight should correspond to the importance of the term for the given corpus. The common weighting scheme is called \emph{Term Frequency Inverse Document Frequency} ($TFIDF$) weighting. An \emph{Inverse Document Frequency} ($IDF$) weight for the dictionary term $k$ is defined as $\log\left( \frac{N}{DF_k} \right)$, where $DF_k$ is the number of documents in the corpus which contain term $k$.  Therefore we can definte a document's $TFIDF$ as
$$ x_{ij}  := \frac{\mbox{term frequency in document } i}{\mbox{inverse document frequency of term } j}
%A document $TFIDF$ vector is its original vector multiplied element-wise by the weights.
%
The $TFIDF$ weighted vector space model document representation corresponds to a map $\phi : \text{text} \rightarrow \RR^n$ defined by:
$$\phi(d)_k = {TF}_k(d) \log\left( \frac{N}{{DF}_k}\right).$$

@@

We will describe the main dataset for building cross-lingual models which is based on Wikipedia. After that we will present three sets of experiments. The first set of experiments
will establish that the hub based approach can deal with language pairs where little or no training data is available. The second set of experiments will compare the main approaches
the we presented on the task of mate retrieval and the task of event linking. We will also examine how different choices of features impact the event linking performance.

@@

To investigate the empirical performance of the low rank approximations we will test the algorithms on a large-scale, real-world multi-lingual dataset that we extracted from Wikipedia by using inter-language links for alignment. This  results in a large number of weakly comparable documents in more than $200$ languages. Wikipedia is a large source of multi-lingual data that is especially important for the languages for which no translation tools, multi-lingual dictionaries as Eurovoc, or strongly aligned multi-lingual corpora as Europarl are available. Documents in different languages are related with so called 'inter-language' links that can be found on the left of the Wikipedia page. The Wikipedia is constantly growing. There are currently 12 Wikipedias with more than 1 million %$10^6$
 articles, $52$ with more than 100k %$10^5$
 articles, $129$ with more than 10k articles, and $236$ with more than $1000$ articles.

@@

We get inter-language link matrix using previously stored redirection links and inter-language links. If inter-language link points to the redirection we replace it with the redirection target link. It turns out that we obtain the matrix $M$ that is not symmetric, consequently the underlying graph is not symmetric. That means that existence of the inter-language link in one way (i.e. English to German) does not guarantee that there is an inter-language link in the reverse direction (German to English). To correct this we transform this matrix to symmetric by computing $M+M^T$ and obtaining an undirected graph. In the rare case that we have multiple links pointing from the document, we pick the first one that we encountered. This matrix enables us to build an alignment across all Wikipedia languages.

@@

 In this subsection, we will investigate the empirical performance of hub CCA approach. We will demonstrate that this approach can be successfully applied even in the case of fully missing alignment information.
 To this purpose, we select a subset of Wikipedia languages containing three major languages, English--\emph{en} (hub language), Spanish--\emph{es}, Russian--\emph{ru}, and five minority (in the sense of Wikipedia sizes) languages, Slovenian--\emph{sl}, Piedmontese--\emph{pms}, Waray-Waray--\emph{war} (all with about 2 million native speakers), Creole--\emph{ht} (8 million native speakers), and Hindi--\emph{hi} (180 million native speakers). For preprocessing, we remove the documents that contain less than 20 different words (stubs) and remove words occurring in less than 50 documents as well as the top 100 most frequent words (in each language separately). We represent the documents as normalized TFIDF\cite{Salton88term-weightingapproaches} weighted vectors.
 %These are languages with Wikipedia article sizes comparable to the minority languages.  (since the number of speakers does not directly correlate with the number of Wikipedia articles). The prime hub candidate is the English %language which is well aligned with all other Wikipedia languages, although the alignment quality varies quite a bit. Furthermore, we remove the documents that contain less than 20 different words and remove words that are too %infrequent as well as the top 100 most frequent words in the vocabularies.
 Although the English language is well aligned with all Wikipedia languages, we must note that quality of alignment varies quite a bit, especially in the case of small Wikipedias.
 Furthermore, we call the document consisting of less than 20 different words, a stub. This documents are typically garbage, the titles of the columns in the table, remains of the parsing process, or Wikipedia articles with very little or no information contained in one or two sentences.

@@

%AMPR as function of the number of feature vectors
%500 is good, 800 better, 1000 overfitting, noise!
\begin{figure}
\label{pic:AMRR}
%\includegraphics{slika.png}
\centering
\input{retrieval}
\caption{Average of mean reciprocal ranks}
\end{figure}
%Why the hub approach?
%Optimize for language pairs in the system that have the largest number of articles in the system!
%Is star CCA really the best???
%Maybe omit this table?
\begin{table}[h]
\caption{Accuracy of cluster linking with 500 - 800 - 1000 features obtained from different cross-lingual similarity algorithms}
\label{table:linkingEvalAlgos}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Models & CA & Precision & Recall \\ \cline{1-4}
  CCA      & 78.2 - 79.6 - 80.3 \% & 76.3 - 78.0 - 80.5 \%  & 81.6 - 82.1 - 79.9 \%  \\ \cline{1-4}
  LSI      & 78.9 - 78.7 - 80.6 \%  & 76.8 - 77.0 - 78.7 \%  & 83.3 - 80.6 - 83.6 \%  \\ \cline{1-4}
 $k$-means & 73.9\phantom{ - 78.7 - 80.6 }\% & 69.5\phantom{ - 78.7 - 80.6 }\%  & 84.6\phantom{ - 78.7 - 80.6 }\%  \\ \cline{1-4}
\end{tabular}
\end{center}
\end{table}

>>>>

We also used a term re-weighting scheme that adjusts for the fact that some words occur more frequently in general. A term weight should correspond to the importance of the term for the given corpus. The common weighting scheme is called \emph{Term Frequency Inverse Document Frequency} ($TFIDF$) weighting. An \emph{Inverse Document Frequency} ($IDF$) weight for the dictionary term $k$ is defined as $\log\left( \frac{N}{DF_k} \right)$, where $DF_k$ is the number of documents in the corpus which contain term $k$.  Therefore we can definte a document's $TFIDF$ as
$$ x_{ij}  := \frac{\mbox{term frequency in document } i}{\mbox{inverse document frequency of term } j}$$
%A document $TFIDF$ vector is its original vector multiplied element-wise by the weights.
%
The $TFIDF$ weighted vector space model document representation corresponds to a map $\phi : \text{text} \rightarrow \RR^n$ defined by:
$$\phi(d)_k = {TF}_k(d) \log\left( \frac{N}{{DF}_k}\right).$$

@@

We will describe the main dataset for building cross-lingual models which is based on Wikipedia. After that we will present two sets of experiments. The first set of experiments
will establish that the hub based approach can deal with language pairs where little or no training data is available. The second set of experiments will compare the main approaches
the we presented on the task of mate retrieval and the task of event linking. We will also examine how different choices of features impact the event linking performance.

@@

To investigate the empirical performance of the low rank approximations we will test the algorithms on a large-scale, real-world multi-lingual dataset that we extracted from Wikipedia by using inter-language links for alignment. This  results in a large number of weakly comparable documents in more than $200$ languages. Wikipedia is a large source of multi-lingual data that is especially important for the languages for which no translation tools, multi-lingual dictionaries as Eurovoc \cite{eurovoc}, or strongly aligned multi-lingual corpora as Europarl \cite{europarl} are available. Documents in different languages are related with so called 'inter-language' links that can be found on the left of the Wikipedia page. The Wikipedia is constantly growing. There are currently 12 Wikipedias with more than 1 million %$10^6$
 articles, $52$ with more than 100k %$10^5$
 articles, $129$ with more than 10k articles, and $236$ with more than $1000$ articles.

@@

We get inter-language link matrix using previously stored redirection links and inter-language links. If inter-language link points to the redirection we replace it with the redirection target link. It turns out that we obtain the matrix $M$ that is not symmetric, consequently the underlying graph is not symmetric. That means that existence of the inter-language link in one way (i.e. English to German) does not guarantee that there is an inter-language link in the reverse direction (German to English). To correct this we transform this matrix to symmetric by computing $M+M^T$ and obtaining an undirected graph. In the rare case that we have multiple links pointing from the document, we pick the first one that we encountered. \todo{Make this more sharp} This matrix enables us to build an alignment across all Wikipedia languages.

@@

 In this subsection, we will investigate the empirical performance of hub CCA approach. We will demonstrate that this approach can be successfully applied even in the case of fully missing alignment information.
 To this purpose, we select a subset of Wikipedia languages containing three major languages, English--\emph{en} (hub language), Spanish--\emph{es}, Russian--\emph{ru}, and five minority (in the sense of Wikipedia sizes) languages, Slovenian--\emph{sl}, Piedmontese--\emph{pms}, Waray-Waray--\emph{war} (all with about 2 million native speakers), Creole--\emph{ht} (8 million native speakers), and Hindi--\emph{hi} (180 million native speakers). For preprocessing, we remove the documents that contain less than 20 different words (stubs) and remove words occurring in less than 50 documents as well as the top 100 most frequent words (in each language separately). We represent the documents as normalized TFIDF\cite{Salton88term-weightingapproaches} weighted vectors.
 %These are languages with Wikipedia article sizes comparable to the minority languages.  (since the number of speakers does not directly correlate with the number of Wikipedia articles). The prime hub candidate is the English %language which is well aligned with all other Wikipedia languages, although the alignment quality varies quite a bit. Furthermore, we remove the documents that contain less than 20 different words and remove words that are too %infrequent as well as the top 100 most frequent words in the vocabularies.
 Although the English language is well aligned with all Wikipedia languages, we must note that quality of alignment varies quite a bit, especially in the case of small Wikipedias.
 Furthermore, we call the document consisting of less than 20 different words, a stub. This documents are typically garbage, the titles of the columns in the table, remains of the parsing process, or Wikipedia articles with very little or no information contained in one or two sentences. \todo{Move this to footnote for stubs}

@@

%AMPR as function of the number of feature vectors
%500 is good, 800 better, 1000 overfitting, noise!
\begin{figure}
%\includegraphics{slika.png}
\centering
\input{retrieval}
\caption{Average of mean reciprocal ranks}
\label{pic:AMRR}
\end{figure}
%Why the hub approach?
%Optimize for language pairs in the system that have the largest number of articles in the system!
%Is star CCA really the best???
%Maybe omit this table?
\begin{table}[h]
\caption{Accuracy of cluster linking with 500 - 800 - 1000 features obtained from different cross-lingual similarity algorithms}
\label{table:linkingEvalAlgos}
\begin{center}
\begin{tabular}{|c|c|c|c|}
  \hline
  \cline{1-4}
  Models & CA & Precision & Recall \\ \cline{1-4}
  CCA      & 78.2 - 79.6 - 80.3 \% & 76.3 - 78.0 - 80.5 \%  & 81.6 - 82.1 - 79.9 \%  \\ \cline{1-4}
  LSI      & 78.9 - 78.7 - 80.6 \%  & 76.8 - 77.0 - 78.7 \%  & 83.3 - 80.6 - 83.6 \%  \\ \cline{1-4}
 $k$-means & 73.9\phantom{ - 78.7 - 80.6 }\% & 69.5\phantom{ - 78.7 - 80.6 }\%  & 84.6\phantom{ - 78.7 - 80.6 }\%  \\ \cline{1-4}
\end{tabular}
\end{center}
\end{table}
